---
title: "Price prediction"
output:
  html_document:
    df_print: paged
date: "2022-12-02"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading packages
```{r}
library(psych)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(stargazer)
library(MASS)
library(caret)
```


## Data importation

```{r}
mydata <- read.csv("C:/Users/dhshe/OneDrive/Desktop/eda/eda/properties_data.csv")
head(mydata)
attach(mydata)
```

# Data cleaning

## Checking for mising values
```{r}
sum(is.na(mydata))
```
## Checking for outliers

```{r}
ggplot(mydata,
aes(x = price)) +
geom_boxplot() +
labs(title = "Apartments price")
```


## Removing outliers

```{r}
#find Q1, Q3, and interquartile range for values in column A
Q1 <- quantile(mydata$price, .25)
Q3 <- quantile(mydata$price, .75)
IQR <- IQR(mydata$price)

#only keep rows in dataframe that have values within 1.5*IQR of Q1 and Q3
data <- subset(mydata, mydata$price> (Q1 - 1.5*IQR) & mydata$price< (Q3 + 1.5*IQR))
attach(data)
```


# Summary statistics

## Selecting numeric variables alone
```{r}
df<-data.frame(price, size_in_sqft, price_per_sqft, no_of_bedrooms, no_of_bathrooms)
```

## Summarize numeric variables
```{r}
describe(df)
```

## Histogram of price

```{r}
ggplot(data,
       aes(x = price)) +
    geom_histogram(bins=50, fill="blue") +
    labs(title = "Histogram showing apartments price")
```

## Price by quality of the apartments
```{r}
ggplot(data,
aes(x = quality,
y = price)) +
geom_boxplot() +
labs(title = "Price by quality of the apartments")
```
## Scatter plots
```{r}
library(GGally)
ggpairs(df)
```

## Histogram of the apartments size in sqft
```{r}
ggplot(data, aes(x = size_in_sqft)) +
geom_histogram(fill = "cornflowerblue",
color = "white") +
labs(title = "Size of the apartment histogram")
```
## Histogram showing distribution of price per sqft
```{r}
ggplot(data, aes(x = price_per_sqft)) +
geom_histogram(fill = "cornflowerblue",
color = "white") +
labs(title = "Price per sqft histogram")
```
## Bar chart showing the distribution of apartments based on quality
```{r}
ggplot(data, aes(x=quality))+geom_bar(fill="blue")+
  labs(title="Quality of apartments")
```

## Creating training and validation set

```{r}
set.seed(23)
trainindex<-createDataPartition(data$price, p =.70, list = FALSE)
train<-data[trainindex,]
validation <- data[-trainindex,]
```



## Modeling

## Stepwise regression
```{r}
#define intercept-only model
intercept_only <- lm(price ~ 1, data=train)

#define model with all predictors
all <- lm(price ~ ., data=train)

#perform backward stepwise regression
backward <- step(all, direction='backward', scope=formula(all), trace=0)

# intercept
backward$coefficients
```

## Linear model picking up variables selected by stepwise model
```{r}
model1<-lm(price~size_in_sqft+price_per_sqft+no_of_bedrooms+covered_parking+kitchen_appliances+lobby_in_building+networked+private_garden+shared_spa+view_of_landmark, data=train)
```

```{r}
summary(model1)
```
## find optimal lambda for Box-Cox transformation 

```{r}
bc <- boxcox(price~size_in_sqft+price_per_sqft+no_of_bedrooms+covered_parking+kitchen_appliances+lobby_in_building+networked+private_garden+shared_spa+view_of_landmark, data=train)
(lambda <- bc$x[which.max(bc$y)])
```


#fit new linear regression model using the Box-Cox transformation

```{r}
new_model <- lm(((price^lambda-1)/lambda) ~ size_in_sqft+price_per_sqft+no_of_bedrooms+covered_parking+kitchen_appliances+lobby_in_building+networked+private_garden+shared_spa+view_of_landmark, data=train)
```

# Summary of the new model

```{r}
summary(new_model)
```

## Prediction and checking accuracy of the model
```{r}
prediction1<-predict(model1, validation)
prediction2<-predict(new_model, validation)

# Converting prediction two back to normal number
prediction3<-(1+lambda*prediction2)^(1/lambda)

attach(validation)
data1<-cbind(price, prediction1, prediction3)
head(data1)

# Calculating RSME to determine the better model
RSME_MODEL1<-sqrt(sum((price-prediction1)^2/532))
RSME_MODEl2<-sqrt(sum((price-prediction3)^2/532))
data.frame(RSME_MODEL1, RSME_MODEl2)
```

